### 2021-06-11 paper/appendix_data.tex
Update figurenotes and small stuff.

Our other two types of contacts, households and schools, get their assortativity by
construction. Schools are groups where the same children of the mostly same age group and
county meet with teachers every day. Household composition follows directly from the
German microcensus data we use to construct our synthetic population.\comment[id=HM]{So
no work in here? How about school? Can we add a column with marginals and another one
with work / (school?), too?}\comment[id=K]{I hope adding work and writing about school and
households addresses all questions. I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}

@@

\subsection{Policies}
\label{subsec:policies}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../figures/results/figures/data/seasonality}
\caption{Seasonality for Other Contacts and Other Contact Models}
\label{fig:seasonality}
\floatfoot{\noindent}
\end{figure}

>>>>

Our other two types of contacts, households and schools, get their assortativity by
construction. Schools are groups where the same children of the mostly same age group and
county meet with teachers every day. Household composition follows directly from the
German microcensus data we use to construct our synthetic population.\comment[id=HM]{So
no work in here? How about school? Can we add a column with marginals and another one
with work / (school?), too?}\comment[id=K]{I hope adding work and writing about school and
households addresses all questions. I think marginals would be confusing because they may not add up to one given that age groups have different group sizes.}\comment[id=HM]{See previous comment. My point is that total number of contacts rather than shares are much more informative in the first place, then different group sizes are no concern anymore.}

@@

\subsection{Policies and Seasonality}
\label{subsec:policies_seasonality}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../figures/results/figures/data/seasonality}
    \caption{Seasonality by Type of Contact}
    \label{fig:seasonality}
    \floatfoot{\noindent}
\end{figure}

### 2021-06-11 paper/report.tex
Update figurenotes and small stuff.

Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
on many dimensions. The development of highly effective vaccines holds the promise of
containment in the medium term. However, most countries find themselves many
months---and often years---away from reaching vaccination-induced herd immunity
\citep{Swaminathan2021}.\comment[id=HM]{Need to find some more scientific citation
before submission} In the meantime, it is of utmost importance to employ an effective
mix of strategies for containing the virus. The most frequent initial response was a set
of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals. While
this has allowed some countries to sustain equilibria with very low infection
numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case
    numbers which is sustained with test-trace-and-isolate policies.}, most have seen large
fluctuations of infection rates. Containment measures have become increasingly diverse
and now include testing, more nuanced NPIs, and contact tracing. Neither these policies'
effect nor the influence of seasonal patterns or more infectious virus strains are well
understood in quantitative terms.

This paper develops a quantitative model incorporating all these factors. The framework
allows to combine a wide variety of data and mechanisms in a timely fashion, making it
useful to predict the effects of various interventions. We apply the model to Germany
and show that rapid testing had the largest impact on the reduction in infections by
almost 80\% during the month of May 2021. We conclude that rapid tests have a large role
to play at least as long as vaccinations have not been offered to an entire population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to
    other approaches in \ref{sec:literature_review}. The model most closely related to ours
    is described in \citet{Hinch2020}.} Each contact between an infectious individual and
somebody susceptible to the disease bears the risk of transmitting the virus. Contacts
occur in the household, at work, at school, or in other settings (leisure activities,
grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
occur at random. Random contacts are typically assortative in age and geographical
location. Empirical applications can take the population structure from census data and
the types and frequency of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to
    multiple data sets on contact types and frequencies covering countries from four
    continents.}

@@

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
        \caption{{Model description}}
        \label{fig:broad_model_description}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
        \vskip4ex

@@

        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
        \caption{{Model for antigen tests}}
        \label{fig:antigen_tests}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
        \caption{{Translating all infections to recorded ones}}
        \label{fig:model_for_official_cases}
    \end{subfigure}

@@

    \floatfoot{\noindent Note: ...}
\end{figure}

In our model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death as for example also modeled in \cite{Grimm2021}.
The probabilities of transitioning between these states depend on age; their duration is
random within intervals calibrated to medical literature (for a detailed description see
Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
independent of age \citep{Jones2021}.

@@

The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_index} the Oxford Response
Stringency Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical
interventions. We transform the index so that lower values represent higher levels of
restrictions. A value of zero means all measures incorporated in the index are turned
on. The value 1 represents the situation in mid-September, with restrictions on
gatherings and public events, masking requirements, but open schools and workplaces. In
the seven weeks between mid September and early November, cases increased by a factor of
10. Restrictions were somewhat tightened in mid-October and again in early November. New
infections remained constant throughout November, before rising again in December, which
prompted the most stringent lockdown to this date. Schools and daycare centers were
closed again, so were customer-facing businesses except for grocery and drug stores.
From the peak of the second wave just before Christmas until the trough in mid-February,
newly detected cases decreased by almost three quarters. The third wave in the spring of
2021 is associated with the B.1.1.7 strain, which became dominant in March. See
Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the public. There
were many changes in details of regulations afterwards, but they did not change the
stringency index.

@@

        \caption{{Stringency of NPIs and infectious contacts}}
        \label{fig:stringency_index}
    \end{subfigure}

@@

    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
        by geographical region. Also more disaggregated data.

        Sources: ...}

@@

The effects of various mechanisms can be disentangled due to the distinct temporal
variation in the drivers of the pandemic. Next to the stringency index, the three lines
in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene
regulations, and seasonality evolved since early September for each of the three broad
contact networks. For example, a value of 0.75 for the work multiplier means that if the
environment was the same as in September (levels of infection rates, no rapid tests or
vaccinations, only the wildtype virus present), infections at the workplace would be
reduced by 25\%. The lines show the product of the effect of contact reductions,
increased hygiene regulations, and seasonality. See Appendix~\ref{subsec:policies} for
separate plots of the three factors. Two aspects are particularly interesting. First, all
lines broadly follow the stringency index and they would do so even more if we left out
seasonality and school vacations (roughly the last two weeks of October, two weeks each
around Christmas and Easter, and some days in late May). Second, the most stringent
regulations are associated with the period of strong decreases in new infections between
late December 2020 and mid-February 2021. The reversal of the trend is associated with he
spread of the B.1.1.7 variant. The steep drop in recorded cases during May 2021 is
associated with at least weekly rapid tests to around 42~percent of the population, a
vaccination rate that rose from 28\% to 43\%, and mostly seasonality impacting a fall in
the relative infectiousness of contacts outside of work and school.

In order to better understand the contributions  of rapid tests, vaccinations, and of
seasonality on the evolution of infections in 2021,
Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always the
same as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the
model fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario
without any of the three factors (red line), and three scenarios turning these factors
off one by one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
values to decompose the difference in total infections between the scenario without any
of the three factors and our main specification.

@@

    \floatfoot{\noindent Note: All aggregates; See S.XXX for statistics by age group and
        by geographical region.

        The decomposition is based on Shapley values where the individual contribution
        of a channel is its average contribution over different sizes of coalitions
        (combinations with other channels). The individual contribution to a coalition
        is the difference between the effect size of the coalition with the particular
        channel and without.}
\end{figure}

>>>>

Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
on many dimensions. The development of highly effective vaccines holds the promise of
containment in the medium term. However, most countries find themselves many
months---and often years---away from reaching vaccination-induced herd immunity
\citep{Swaminathan2021}.\comment[id=HM]{Need to find some more scientific citation
before submission} In the meantime, it is of utmost importance to employ an effective
mix of strategies for containing the virus. The most frequent initial response was a set
of non-pharmaceutical interventions (NPIs) to reduce contacts between individuals. While
this has allowed some countries to sustain equilibria with very low infection
numbers\footnote{See \citet{Contreras2021} for a theoretical equilibrium at low case
numbers which is sustained with test-trace-and-isolate policies.}, most have seen large
fluctuations of infection rates over time. Containment measures have become increasingly
diverse and now include testing, more nuanced NPIs, and contact tracing. Neither these
policies' effect nor the influence of seasonal patterns or more infectious virus strains
are well understood in quantitative terms.

This paper develops a quantitative model incorporating all these factors. The framework
allows to combine a wide variety of data and mechanisms in a timely fashion, making it
useful to predict the effects of various interventions. We apply the model to Germany
and show that rapid testing had the largest impact on the reduction in infections by
almost 80\% during the month of May 2021. We conclude that rapid tests have a large role
to play for at least as long as vaccinations have not been offered to an entire
population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:broad_model_description}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
model most closely related to ours is described in \citet{Hinch2020}.} Each contact
between an infectious individual and somebody susceptible to the disease bears the risk
of transmitting the virus. Contacts occur in the household, at work, at school, or in
other settings (leisure activities, grocery shopping, medical appointments, etc.). Some
contacts recur regularly, others occur at random. Random contacts are typically
assortative in age and geographical location. Empirical applications can take the
population structure from census data and the types and frequency of contacts from diary
data measuring contacts before the pandemic \citep[e.g.][]{Mossong2008,Hoang2019}.

@@

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
        \caption{{Model description}}
        \label{fig:broad_model_description}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
        \vskip5ex

@@

        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
        \caption{{Model for antigen tests}}
        \label{fig:antigen_tests}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-right}
        \vskip5ex

        \caption{{Translating all infections to recorded ones}}
        \label{fig:model_for_official_cases}
    \end{subfigure}

@@

    \floatfoot{\noindent \textit{Note:}
    A % formal 
    description of the model can be found in Supplementary Material~\ref{sec:model}.
    Figure~\ref{fig:broad_model_description} shows the influence on an agent's contacts
    to other agents. Demographic characteristics set the baseline number of contacts in
    different networks. She may reduce the number of contacts due to NPIs, showing
    symptoms, or testing positively for SARS-CoV-2. Infections may occur when a
    susceptible agent meets an infectious agent; the probability depends on the type of
    contact, on seasonality, and on NPIs. If infected, the infection progresses as
    depicted in Figure~\ref{fig:disease_progression}. If rapid tests are avaible,
    agents' demand is modeled as in Figure~\ref{fig:antigen_tests}. All pupils are
    attending school are tested, other reasons trigger a test only for a fraction of
    individuals. Figure~\ref{fig:model_for_official_cases} shows the model of
    translating all infections in the simulated data to age-specific recorded
    infections. The model uses data on the aggregate share of recorded cases, on the
    share of positive PCR tests triggered by symptoms, and on the false positive rate of
    rapid tests. The lower part of the graph is relevant only for periods where rapid
    tests are available.}
\end{figure}

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death as for example also modeled in \cite{Grimm2021}.
The probabilities of transitioning between these states depend on age; their duration is
random within intervals calibrated to medical literature (for a detailed description see
Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
independent of age \citep{Jones2021}.

@@

The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the Oxford Response
Stringency Index \citep{Hale2020}, which tracks the tightness of non-pharmaceutical
interventions. We transform the index so that lower values represent higher levels of
restrictions. A value of zero means all measures incorporated in the index are turned
on. The value 1 represents the situation in mid-September, with restrictions on
gatherings and public events, masking requirements, but open schools and workplaces. In
the seven weeks between mid September and early November, cases increased by a factor of
10. Restrictions were somewhat tightened in mid-October and again in early November. New
infections remained constant throughout November, before rising again in December, which
prompted the most stringent lockdown to this date. Schools and daycare centers were
closed again, so were customer-facing businesses except for grocery and drug stores.
From the peak of the second wave just before Christmas until the trough in mid-February,
newly detected cases decreased by almost three quarters. The third wave in the spring of
2021 is associated with the B.1.1.7 strain, which became dominant in March. See
Figure~\ref{fig:share_b117}. In early March, some NPIs were being relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the public. There
were many changes in details of regulations afterwards, but they did not change the
stringency index.

@@

        \caption{{Stringency of NPIs and infectious contacts}}
        \label{fig:stringency_infectious_contacts}
    \end{subfigure}

@@

    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
    Material~\ref{sec:data_and_parameters}. Age- and region-specific analogues to
    Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
    \ref{subsec:new_cases_fit}. For legibility reasons, all lines in
    Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
    Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that a
    value of 1 is refers to the situation at the start of our sample period and 0 means
    that all NPIs included in the index are turned on. The other lines in
    Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect of
    contact reductions, increased hygiene regulations, and seasonality. See
    Appendix~\ref{subsec:policies_seasonality} for separate plots of the three factors
    by contact type.}
\end{figure}

@@

The effects of various mechanisms can be disentangled due to the distinct temporal
variation in the drivers of the pandemic. Next to the stringency index, the three lines
in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions, increased hygiene
regulations, and seasonality evolved since early September for each of the three broad
contact networks. For example, a value of 0.75 for the work multiplier means that if the
environment was the same as in September (levels of infection rates, no rapid tests or
vaccinations, only the wildtype virus present), infections at the workplace would be
reduced by 25\%. The lines show the product of the effect of contact reductions,
increased hygiene regulations, and seasonality. Two aspects are particularly
interesting. First, all lines broadly follow the stringency index and they would do so
even more if we left out seasonality and school vacations (roughly the last two weeks of
October, two weeks each around Christmas and Easter, and some days in late May). Second,
the most stringent regulations are associated with the period of strong decreases in new
infections between late December 2020 and mid-February 2021. The reversal of the trend
is associated with he spread of the B.1.1.7 variant. The steep drop in recorded cases
during May 2021 is associated with at least weekly rapid tests to around 42~percent of
the population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
impacting a fall in the relative infectiousness of contacts outside of work and school.

In order to better understand the contributions  of rapid tests, vaccinations, and of
seasonality on the evolution of infections in 2021,
Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always the
same as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the
model fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario
without any of the three factors (red line), and three scenarios turning these factors
off one by one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
values\comment[id=HM]{Need reference} to decompose the difference in total infections
between the scenario without any of the three factors and our main specification.

@@

    \floatfoot{\noindent \textit{Note:} The blue line in
        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
        variant is introduced in the same way; vaccinations, rapid tests, and
        seasonality remain at their January levels. The other scenarios turn these three
        factors on one-by-one. The decompositions in
        Figures~\ref{fig:2021_scenarios_decomposition} and
        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values
        where the individual contribution of a channel is its average contribution over
        different sizes of coalitions (combinations with other channels). The individual
        contribution to a coalition is the difference between the effect size of the
        coalition with the particular channel and without.}
\end{figure}
